---
title: "Homework 4: Machine Learning"
author: "Patrick Hofbauer"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)

```

# The Bechdel Test

<https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/>

The [Bechdel test](https://bechdeltest.com) is a way to assess how women are depicted in Hollywood movies. In order for a movie to pass the test:

1.  It has to have at least two [named] women in it
2.  Who talk to each other
3.  About something besides a man

There is a nice article and analysis you can find here <https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/> We have a sample of 1394 movies and we want to fit a model to predict whether a film passes the test or not.

```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)


```

How many films fail/pass the test, both as a number and as a %?

```{r}
#counting pass/fail and getting percentages
bechdel %>% 
  
  group_by(test) %>% 
  
  summarise(count = n()) %>% 
  
  mutate(total = sum(count), prop = count/total)
```

772 films have failed the Bechdel test and 622 films have passed it. The represents a failure rate of 55.4%.

## Movie scores

```{r}

#plotting the IMDB scores
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test #grouping by pass/fail
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) + #changing the colors of the groups
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light() #changing the theme of the graph
```

# Split the data

```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.

```{r}
bechdel_train %>% group_by(test) %>% summarise(count = n()) %>% mutate(total = sum(count), prop = count/total)

bechdel_test %>% group_by(test) %>% summarise(count = n()) %>% mutate(total = sum(count), prop = count/total)
```

## Feature exploration

## Any outliers?

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

```

In all selected categories the variables are on average higher if the film fails the bechdel test. There are a few outliers, for example in film budget one of the films that failed the Bechdel test, had a budget over 40m. In IMDB ratings, a few films that failed the Bechdel test had very low ratings and should be classified as outliers. There are also films in both test categories that grossed an unusual amount, both domesticlly and internationally.


## Scatterplot - Correlation Matrix

Write a paragraph discussing the output of the following

```{r, warning=FALSE, message=FALSE}

#plotting the correlation matrix of the dataframe 
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% #selecting the required variables
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()
```

## Categorical variables

Write a paragraph discussing the output of the following

```{r, warning=FALSE}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% #.groups = "drop.last"
  mutate(prop = n/sum(n))
```

The tables analyses the passing of the Bechdel test based on film genres. Sci-Fi and Documentary films have a 100% failure rate. Action, Mystery and Horror movies are the next categories to have the highest failure rates. Only Musicals and Thrillers have a 100% pass rate.

```{r, warning=FALSE}
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
```

The table assess the bechdel test results of films based on the age restriction rating. It seems that films with a tighter restriction have a higher failure rate in the bechdel test. There is however a slight exceptions for films with a G rating. The highest failure rate is for films with a NC-17 rating andf the lowest is for films with a PG-13 rating.

# Train first models. `test ~ metascore + imdb_rating`

```{r}
#setting up the base model for logistic regression 
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% #choosing the engine that we run the regression with
  set_mode("classification") #choosing the classification mode

lr_mod

#setting up the base model for a decision tree
tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod 
```

```{r}

#fitting the logistic regression model to the required variables
lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

#fitting the decision tree model to the required variables
tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

lr_fit
```

## Logistic regression

```{r}
#testing for prediction matches
lr_fit %>%
  broom::tidy() #converting stat objects into tidy tibbles

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))


```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")


```

## Decision Tree

```{r}

#testing for prediction matches
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return?

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 10, 
                          strata = test)
bechdel_folds
```

The code creates 10 new random data splits (folds) from the training data






CHECKKKKKKKKKK










## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```

## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`

```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)


```

```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like?
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

```

# Build a better training set with `recipes`

## Preprocessing options

-   Encode categorical predictors
-   Center and scale variables
-   Handle class imbalance
-   Impute missing data
-   Perform dimensionality reduction
-   ... ...

## To build a recipe

1.  Start the `recipe()`
2.  Define the variables involved
3.  Describe **prep**rocessing [step-by-step]

## Collapse Some Categorical Levels

Do we have any `genre` with few observations? Assign genres that have less than 3% to a new category 'Other'

```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```

```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 5% will be in a category called 'Other'
    step_other(genre, threshold = .03) 
```

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```

## After recipe

```{r}
movie_rec %>% 
  prep() %>% #estimate the parameters required of the recipe
  bake(new_data = bechdel_train) %>% #apply the newly trained recipe
  count(genre, sort = TRUE)
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling

What if there were no films with `rated` NC-17 in the training data?

-   Will the model have a coefficient for `rated` NC-17?
-   What will happen if the test data includes a film with `rated` NC-17?

## `step_novel()`

Adds a catch-all level to a factor for any new values not encountered in model training, which lets R intelligently predict new levels in the test set.

```{r}

movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal_predictors) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal_predictors()) 

```

## `step_zv()`

Intelligently handles zero variance variables (variables that contain only a single value)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes()) 
  
```

## `step_normalize()`

Centers then scales numeric variable (mean = 0, sd = 1)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

```

## `step_corr()`

Removes highly correlated variables

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) %>% 
  step_corr(all_predictors(), threshold = 0.75, method = "spearman") 



movie_rec
```

# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```

# Bundle recipe and model with `workflows`

```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1.  How many models have you specified?
2.  What's the difference between a model specification and a workflow?
3.  Do you need to add a formula (e.g., `test ~ .`) if you have a recipe?

# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification.

```{r}

```

# Deliverables

There is a lot of explanatory text, comments, etc. You do not need these, so delete them and produce a stand-alone document that you could share with someone. Knit the edited and completed R Markdown (Rmd) file as a Word or HTML document (use the "Knit" button at the top of the script editor window) and upload it to Canvas. You must be commiting and pushing your changes to your own Github repo as you go along.

# Details

-   Who did you collaborate with: Nobody
-   Approximately how much time did you spend on this problem set: 1h
-   What, if anything, gave you the most trouble: Descriptive Statistics

**Please seek out help when you need it,** and remember the [15-minute rule](https://dsb2023.netlify.app/syllabus/#the-15-minute-rule){target="_blank"}. You know enough R (and have enough examples of code from class and your readings) to be able to do this. If you get stuck, ask for help from others, post a question on Slack-- and remember that I am here to help too!

> As a true test to yourself, do you understand the code you submitted and are you able to explain it to someone else?

YES


# Rubric

13/13: Problem set is 100% completed. Every question was attempted and answered, and most answers are correct. Code is well-documented (both self-documented and with additional comments as necessary). Used tidyverse, instead of base R. Graphs and tables are properly labelled. Analysis is clear and easy to follow, either because graphs are labeled clearly or you've written additional text to describe how you interpret the output. Multiple Github commits. Work is exceptional. I will not assign these often.

8/13: Problem set is 60--80% complete and most answers are correct. This is the expected level of performance. Solid effort. Hits all the elements. No clear mistakes. Easy to follow (both the code and the output). A few Github commits.

5/13: Problem set is less than 60% complete and/or most answers are incorrect. This indicates that you need to improve next time. I will hopefully not assign these often. Displays minimal effort. Doesn't complete all components. Code is poorly written and not documented. Uses the same type of plot for each graph, or doesn't use plots appropriate for the variables being analyzed. No Github commits.
